---
format: gfm
---

These look good, and I'm glad you are highlighting the distributions of the two indices. The main thing is your charts are different ways to show the same two variables, so they become redundant. The histogram and the density plot are different ways of showing essentially the same thing---they're both good, but you should pick one, and then do something different for another chart. You had talked earlier about joining the EJ data with the ACS data, and I think that's still a great idea. There are some examples of that in the troubleshooting notes, but I can also help you with specific steps if that's something you want to do. Some more specific suggestions:

## Debugging

In your final draft notebook, when you draw the histogram you get a warning about a bunch of values being removed. You might expect a handful of missing values (in your EDA you checked the number of NAs, excellent call!), but this says there are 1,462. That's a red flag, so you want to go back and look at the structure of the data you're plotting (ej_spread). In lines 33 and 34 where you call `pivot_wider`, you're getting a bunch of NAs:

```{r}
#| message: false
library(justviz)
library(dplyr)
library(ggplot2)

ej_subset <- ej_natl |>
  filter(indicator %in% c("releases_to_air", "wastewater"))

head(ej_subset)

ej_spread1 <- ej_subset |>
  tidyr::pivot_wider(names_from = indicator, values_from = value_ptile)

head(ej_spread1)
sum(is.na(ej_spread1))
```

In ej_spread, your tracts are duplicated (2 rows each when you would expect 1 row each) and your two indicator columns alternate between NAs. The reason for that is `pivot_wider` doesn't know to drop the d2 and d5 columns. It expects there to be one or more identifier columns (the `id_cols` argument), and in order to avoid dropping data when you don't expect it, that defaults to all the columns in the data frame that you haven't otherwise specified. In this case, since you're only concerned with the value_ptile values, you want to specify an ID, which is tract:

```{r}
ej_spread2 <- ej_subset |>
  tidyr::pivot_wider(id_cols = tract, names_from = indicator, values_from = value_ptile)

head(ej_spread2)
sum(is.na(ej_spread2))
```

There are still a few NAs, but more in the range you'd expect.

The other red flag is that your scatterplot has so few points, and it's for the same reason, that most of your pairs of values per tract got lost. You should expect it to look more like this, with a warning about a few NAs being removed:


```{r}
ggplot(ej_spread2, aes(x = releases_to_air, y = wastewater)) +
  geom_point(alpha = 0.5)
```

However, I see you also want to color by d2 percentile. Since each tract actually has 2 d2 values (one for releases, one for wastewater) you have to make a decision here. Think about what the d2 values bring into the mix that's not already in the scatterplot: they scale by low-income rate and percent people of color, and that scaling is the same for the two risk factors. Both low-income rate and percent POC are available in the ACS data (since the race/ethnicity groups are percentages, you can just do 1 - white to get percent non-white), so if you join the EJ data with ACS, you could color by one (or both, and then you've knocked out 2 of your 3 charts). Here's an example that is _extremely similar_ but a different ACS variable so you can still figure out how you want to approach yours:


```{r}
ej_x_acs <- ej_spread2 |>
  inner_join(acs, by = c("tract" = "name"))

ggplot(ej_x_acs, aes(x = releases_to_air, y = wastewater, color = foreign_born)) +
  geom_point(alpha = 0.5)
```

## Scatterplot

Aside from the debugging, give some information on what these values mean (esp since your goal was to have titles and labels match the data well). Specify the year (2022), unit (Maryland tract), and source (EPA EJSCREEN). Give information on how these values were calculated: you can fit some of that in the subtitle (`labs(subtitle = "Nationally ranked percentiles of blah blah blah")`), or fit more detail into a caption with the source (`labs(caption = "blah blah")`). You also don't need to say that it's a scatterplot in the title; since your charts are a bit more technical, you can expect that your audience will understand that. You could just say something like "Risk of toxins released to air versus proximity to wastewater treatment".

## Histogram and density

Like I said, you should pick one of the two of these. On the histogram, notice that you didn't get a legend---the reason for that is you don't have anything fill-related in your `aes`. You kinda hacked it into your density chart. For either of these, what you want to do is use the long-shaped data instead (`ej_subset`), and assign the indicator to fill (for the histogram) or color and/or fill (for the density).

I like the idea of annotating with a dashed line like on the density plot, although that information is already on the axis. Drawing a line for the mean and/or median value is often helpful for these types of distribution plots though. You could do something like this (technically you'd want to calculate weighted means or medians, weighted by population, but this simplified version is fine for this project):


```{r}
# save the colors you made into a vector to reuse
ej_pal <- c(releases_to_air = "#990000", wastewater = "#034e7b")
ej_avgs <- ej_subset |>
  group_by(indicator) |>
  summarise(med_value_ptile = median(value_ptile, na.rm = TRUE))

ej_avgs

ggplot(ej_subset, aes(x = value_ptile, fill = indicator)) +
  geom_histogram(alpha = 0.3, 
                 position = position_identity(), # use identity so they don't get stacked
                 binwidth = 3) + # make sure to adjust the binwidth or number of bins!
  geom_vline(aes(xintercept = med_value_ptile, color = indicator), data = ej_avgs) +
  scale_fill_manual(values = ej_pal) +
  scale_color_manual(values = ej_pal)

```

And then make a note in the subtitle stating that the lines are medians. That's just one idea though.